\section{Discussion and Future Work}
\label{sec:discussion}

This pilot establishes two outputs: a feasibility profile for the generation-and-audit pipeline, and preliminary proxy-based directional patterns to prioritise for human evaluation. On feasibility, roadway-marking and maintenance interventions realise reliably across cities, while lighting repair remains the clearest bottleneck. No meaningful relationship is observed between baseline safety and valid lever count, suggesting that multi-lever richness depends more on scene geometry than starting safety level. On directionality, Mobility Infrastructure yields the largest retained auxiliary shifts, though it remains unclear whether this reflects human perceptual sensitivity, scorer priors, or edit realisability.

Environmental Amenity is more heterogeneous. Localised greenery addition is strongly positive, whereas lighting repair and canopy management are negative on average despite being theory-grounded cues. This contrast suggests that the auxiliary scorer may encode a simple “more green = safer” prior rather than a more nuanced sensitivity to sightlines or surveillance, and that family-level aggregation can obscure important lever-specific differences.

\noindent\textbf{On the model-based evaluation proxy.} The scorer $f_a$ has not been validated on edited images, so $\Delta_i$ should not be read as a calibrated effect size. However, for monotone interventions (road markings appear, graffiti is removed, greenery is added), the directional signal is grounded in large-scale crowdsourced pairwise preference over this class of scene variation, making the sign and relative magnitude of $\Delta_{\mathrm{aux}}$ a useful prioritisation signal for selecting which levers and scenes to send to human evaluation.


\noindent\textbf{XAI contribution.} Beyond urban perception, the methodological contribution is a stronger standard for visual explanation: the proposed change must be semantically coherent and generatively feasible before it enters the attribution result, providing a reusable scaffold for interventional XAI in other perceptual domains.


\noindent\textbf{Future directions and human endpoint.}
The present pipeline is intentionally minimal, prompt-only generation, a prompted VLM critic, and an auxiliary proxy scorer are used here to establish a reproducible baseline rather than a fully validated explanation system. Four directions follow naturally from the
pipeline structure. \emph{(i) Planner tuning:} learning to predict lever
viability from scene features, using the validity and
$\Delta_{\mathrm{aux}}$ outcomes reported here as supervision, so that
the $K$-candidate budget is spent on levers most likely to be both
realisable and promising under the auxiliary proxy. \emph{(ii) Generator alignment:} optimising
for validity-gate criteria via reward-from-feedback or mask-constrained
generation; the lighting-repair failure mode (valid rate 0.30) is the
clearest target, with per-family valid rates as the benchmark metric.
\emph{(iii) Critic calibration:} aligning the VLM critic with human
raters and ultimately learning a critic that predicts 2AFC preference
directly, reducing the gap between the auxiliary proxy and the human
endpoint. \emph{(iv) End-to-end optimisation:} jointly training planner,
generator, and critic once each component is independently validated.
The ground-truth estimand throughout is pairwise human safety ratings on passing edits, collected via a randomised 2AFC study. Until then,
validity should be read as an eligibility control on image generation
and auxiliary shifts as a prioritisation signal for which lever
families and cities to send to human evaluation first.
