\section{Discussion and Future Work}
\label{sec:discussion}

\noindent\textbf{Claim scope.}
We frame results as \emph{valid and directionally promising} levers, not
independent causal levers. Additivity of $\delta_i$ and $\delta_j$ is
untested; joint effects require factorial designs left for future work.
Because prompt-only editing can introduce residual texture changes
outside the intended support~\cite{bazyleva2025xedit}, locality should
be read as a pragmatic validity screen rather than pixel-accurate
isolation, and we position the work as interventional attribution rather
than structural causal inference.

\noindent\textbf{Vocabulary scope and grounding bias.}
Results are measured only over the bounded intervention vocabulary
(\S\ref{sec:method}). Cues outside that scope, or too structurally
complex for prompt-only editing, are excluded; $|V(x)|{=}0$ may reflect
this limitation rather than a perception ceiling. Planner grounding may
further favour large, salient elements, so multi-lever counts are likely
underestimates.

\noindent\textbf{What the larger run changes.}
At $N{=}50$, the dominant story is no longer pipeline instability.
After rerunning the previously quota-blocked images, the final dataset
contains no pre-audit generator or planner failures. The main source of
variance is instead lever-specific editability and scorer response:
maintenance and roadway-marking interventions realise reliably, while
lighting repair and canopy-management edits remain harder to realise or
less consistently aligned with auxiliary safety improvement.

\noindent\textbf{Mean effects versus retained positive tail.}
The new cutoff analysis sharpens that point. Family-level mean
$\Delta_{\mathrm{aux}}$ is useful, but it can hide whether a family
produces many modestly positive levers per scene or only a narrow set of
high-impact ones. Mobility Infrastructure remains strong under stricter
cutoffs, while Environmental Amenity looks bimodal: localized greenery
addition drives a substantial positive tail, but lighting repair and
tree canopy management dilute the family mean. This is a more useful
screen for downstream human evaluation than raw validity alone, because
it identifies which families continue to supply multiple
$\theta_{\mathrm{aux}}$-effective candidates once neutral edits are
removed.

\noindent\textbf{Critic calibration.}
The VLM critic (GPT-5.2) has not been calibrated against human raters.
Even with 177 accepted edits, a modest false-positive or false-negative
rate still matters because the paper's claim depends on accepting only
same-place, local, and plausible single-lever edits. An inter-rater study
(human annotators judging the same pool on the four audit criteria,
reporting Cohen's $\kappa$ per criterion) is a priority next step.

\noindent\textbf{Baseline score and realisability.}
Whether low-baseline scenes admit more valid levers than high-baseline
scenes remains an open question. The larger $N{=}50$ run still shows no
meaningful monotonic relationship (Spearman $\rho{=}0.095$,
$p{=}0.511$), now without the confound of missing candidate generation.
This suggests that valid lever abundance is not simply a function of the
starting safety score; scene geometry, visible editable supports, and
the generator's competence on specific lever families likely matter
more.

\noindent\textbf{City and family heterogeneity.}
The cleaned rerun also makes the cross-city picture interpretable.
Amsterdam and Santiago show the highest valid rates, while Singapore and
Abuja are weaker on average, and mean auxiliary movement is strongest in
Santiago and San Francisco. These differences do not yet justify strong
substantive city-level claims, but they do suggest that both
realisability and perceived effect are context-sensitive rather than
uniform across urban settings.

The thresholded view makes the heterogeneity more concrete. San
Francisco and Santiago contribute the highest average number of
$\theta_{\mathrm{aux}}$-effective edits, and Amsterdam stays competitive
once the cutoff is raised, whereas Abuja falls off more quickly. That
pattern suggests the practical ceiling is not only whether a scene
admits valid levers, but whether it admits \emph{multiple} visibly
grounded edits that survive increasingly strict directional thresholds.

\noindent\textbf{Human evaluation.}
The present paper reports automated auditing and auxiliary-scorer
diagnostics only. The planned endpoint is a two-stage human study: an
E1 validity check (same-place, locality, realism, plausibility)
followed by a randomised 2AFC safety comparison on E1-passing edits.
Until then, results should be read as validating the intervention
protocol and its failure modes, not as definitive estimates of which
levers shift perceived safety.
